---
title: "p8105_hw2_jw4489"
output: github_document
date: "2023-10-04"
---

```{r}
library(tidyverse)
```


##Problem 1

```{r}
pols_month_df = 
  read_csv("p1data/pols-month.csv") |> 
  janitor::clean_names() |> 
  separate(mon, into = c("year", "month", "day"), sep = "-") |> 
  mutate(
    month = month.name[as.numeric(month)],
    year = as.numeric(year),
    president = ifelse(
      prez_gop > prez_dem, "gop", "dem"
    )
  ) |> 
  select(-day, -prez_gop, -prez_dem)

pols_month_df
```

```{r}
snp_df = 
  read_csv("p1data/snp.csv") |>
  janitor::clean_names() |> 
  mutate(
    date = format(as.Date(date, format = "%m/%d/%y"), "%Y/%m/%d")
  ) |> 
  separate(date, into = c("year", "month", "day"), sep = "/") |> 
  mutate(
    month = month.name[as.numeric(month)],
    year = as.numeric(year),
    year = ifelse(year > 2023, year - 100, year)
  ) |> 
  select(-day)

snp_df
```

```{r}
unemployment_df = 
  read_csv("p1data/unemployment.csv") |> 
  janitor::clean_names() |> 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment"
  ) |> 
  mutate(
    month = month.name[match(month, tolower(month.abb))],
    year = as.numeric(year)
  )

unemployment_df
```

```{r}
merged_df = left_join(pols_month_df,snp_df, by = c("year", "month")) |> 
  left_join(unemployment_df, by = c("year", "month"))

merged_df
```

The `pols-month` dataset contains `r nrow(pols_month_df)` and `r ncol(pols_month_df)` columns. It contains various variables including `year`, `month`, `gov_gop`, `sen_gop`, `rep_gop`, `gov_dem`, `sen_dem`, `rep_dem`, `president`. The range is `r pull(pols_month_df, year) |> min()` to `r pull(pols_month_df, year) |> max()`. 

The `snp` dataset contains `r nrow(snp_df)` rows and `r ncol(snp_df)` columns. It contains `year`, `month`, `close`. The range is `r pull(snp_df, year) |> min()` to `r pull(snp_df, year) |> max()`.

The `unemployment` dataset contains `r nrow(unemployment_df)` rows and `r ncol(unemployment_df)` columns. It contains `year`, `month`, `unemployment`. The range is `r pull(unemployment_df, year) |> min()` to `r pull(unemployment_df, year) |> max()`.

The merged dataset `merged_df` contains `r nrow(merged_df)` observations and `r ncol(merged_df)`. It contains various variables including `year`, `month`, `gov_gop`, `sen_gop`, `rep_gop`, `gov_dem`, `sen_dem`, `rep_dem`, `president`, `close`. Columns `close` and `unemployment` have the values of `NA`, meaning that there are missing data. 
