---
title: "p8105_hw2_jw4489"
output: github_document
date: "2023-10-04"
---

```{r}
library(tidyverse)
library(readxl)
```


##Problem 1

```{r}
pols_month_df = 
  read_csv("p1data/pols-month.csv") |> 
  janitor::clean_names() |> 
  separate(mon, into = c("year", "month", "day"), sep = "-") |> 
  mutate(
    month = month.name[as.numeric(month)],
    year = as.numeric(year),
    president = ifelse(
      prez_gop > prez_dem, "gop", "dem"
    )
  ) |> 
  select(-day, -prez_gop, -prez_dem)

pols_month_df
```

```{r}
snp_df = 
  read_csv("p1data/snp.csv") |>
  janitor::clean_names() |> 
  mutate(
    date = format(as.Date(date, format = "%m/%d/%y"), "%Y/%m/%d")
  ) |> 
  separate(date, into = c("year", "month", "day"), sep = "/") |> 
  mutate(
    month = month.name[as.numeric(month)],
    year = as.numeric(year)) |>
  arrange(year, month) |>
  relocate(year, month, everything())

snp_df
```

```{r}
unemployment_df = 
  read_csv("p1data/unemployment.csv") |> 
  janitor::clean_names() |> 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment"
  ) |>
  arrange(year, month)

unemployment_df
```

```{r}
merged_df = left_join(pols_month_df,snp_df, by = c("year", "month")) |> 
  left_join(unemployment_df, by = c("year", "month"))

merged_df
```

The `pols-month` dataset contains `r nrow(pols_month_df)` and `r ncol(pols_month_df)` columns. It contains various variables including `year`, `month`, `gov_gop`, `sen_gop`, `rep_gop`, `gov_dem`, `sen_dem`, `rep_dem`, `president`. The range is `r pull(pols_month_df, year) |> min()` to `r pull(pols_month_df, year) |> max()`. 

The `snp` dataset contains `r nrow(snp_df)` rows and `r ncol(snp_df)` columns. It contains `year`, `month`, `close`. The range is `r pull(snp_df, year) |> min()` to `r pull(snp_df, year) |> max()`.

The `unemployment` dataset contains `r nrow(unemployment_df)` rows and `r ncol(unemployment_df)` columns. It contains `year`, `month`, `unemployment`. The range is `r pull(unemployment_df, year) |> min()` to `r pull(unemployment_df, year) |> max()`.

The merged dataset `merged_df` contains `r nrow(merged_df)` observations and `r ncol(merged_df)`. It contains various variables including `year`, `month`, `gov_gop`, `sen_gop`, `rep_gop`, `gov_dem`, `sen_dem`. 

##Problem 2

```{r}
trash_wheel_df = 
  read_excel("p2data/202309 Trash Wheel Collection Data.xlsx", 
             sheet = "Mr. Trash Wheel") |> 
  janitor::clean_names() |>
  drop_na(dumpster) |> 
  select(dumpster:homes_powered)|> 
  mutate(
    homes_powered = (weight_tons * 500) / 30,
    name = "Mr. Trash Wheel",
    year = as.numeric(year)
  )

trash_wheel_df
```

```{r}
prof_df = 
  read_excel("p2data/202309 Trash Wheel Collection Data.xlsx", 
             sheet = "Professor Trash Wheel") |> 
  janitor::clean_names() |>
  drop_na(dumpster) |> 
  mutate(
    homes_powered = (weight_tons * 500) / 30,
    name = "Professor Trash Wheel",
    year = as.numeric(year)
  )

prof_df
```

```{r}
gwynnda_df = 
  read_excel("p2data/202309 Trash Wheel Collection Data.xlsx", 
             sheet = "Gwynnda Trash Wheel") |> 
  janitor::clean_names() |>
  drop_na(dumpster) |> 
  mutate(
    homes_powered = (weight_tons*500)/30,
    name = "Gwynnda Trash Wheel",
    year = as.numeric(year)
  )

gwynnda_df
```

```{r}
combined_df = bind_rows(trash_wheel_df, prof_df, gwynnda_df) |> 
  select(trash_wheel_name = "name",everything())

combined_df
```

The `Mr. trash wheel` dataset contains `r nrow(trash_wheel_df)` rows and `r ncol(trash_wheel_df)` columns.

The `Professor Trash Wheel` dataset contains `r nrow(prof_df)` rows and `r ncol(prof_df)` columns. For available data, the total weight of trash collected by Professor Trash Wheel is `r sum(prof_df|>pull(weight_tons))` tons.

The `Gwynnda Trash Wheel` dataset contains `r nrow(gwynnda_df)` rows and `r ncol(gwynnda_df)` columns. For available data, the total number of cigarette butts collected by Gwynnda in July of 2021 is `r subset(gwynnda_df, year == 2021 & month == "July") |> pull(cigarette_butts) |> sum()|>as.integer()`.

The combined dataset contains `r nrow(combined_df)` rows and `r ncol(combined_df)` columns. 

The datasets contains various variables, inclusing “month”, “year”, “date”, “weight_tons”, “volume_cubic_yards”, “plastic_bottles”, and “polystyrene”. These variables represent the characteristics of trash collected.




